The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-10-21 paper/ms.tex

\begin{abstract}
We present novel results of radiative transfer calculations to measure
the impact of gas bulk rotation on the morphology of the Lyman
$\alpha$ line in galaxies. We model a galaxy as a sphere with an
homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with
maximum velocities in the range $0-300$ \kms. The neutral
hydrogen optical depths are in the range $\tau_{\rm
H}=10^{5}-10^{7}$. We also consider two kinds of spatial
distribution for the radiation sources with respect to the sphere:
central and homogeneous. We find that the line width and the amount of
Lyman alpha photons escaping in the line center increase with
rotational velocity. higher rotational velocities increase the width
of each peak in the outgoing line profile while it also increases the
amount of Lyman alpha photons escaping in the line. For large
rotational velocities and high Hydrogen optical depths the outgoing
line presents single peak a the line center. This trend is more
pronounced for radiation sources homogeneously distributed in the
sphere. Rotation increases by a factor of $2$ the escape fraction for
systems with homogeneously distributed sources. For radiation sources
located off-center, the line morphology presents a range of single,
double and triple peak lines. This work demonstrates that gas bulk
rotation has a non negligible impact on the morphology of the Lyman
$\alpha$ line.

@@

This paper is paper is structured as follows. In \S
\ref{sec:implementation} we present the implementation of bulk
rotation into the Monte Carlo codes, paying special attention to coordinate
definitions. There we also list the different physical parameters in
the simulated grid of models. In the next section we present the
results of the simulations, with special detail to quantities in the
line that show a clear evolution as a function of the cloud's
rotational velocity. In \S \ref{sec:dicussion} we discuss the
implications of our results in the interpretation of LAEs
observations and further refinements that can be implemented in the
theoretical modelling of the \ly line. In the last section we present
our conclusions.

@@

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
Figure clearly shows that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$ \kms.

>>>>

\begin{abstract}
We present novel results of radiative transfer calculations to measure
the impact of gas bulk rotation on the morphology of the Lyman
$\alpha$ line in galaxies. We model a galaxy as a sphere with an
homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with
maximum velocities in the range $0-300$ \kms. The neutral
hydrogen optical depths are in the range $\tau_{\rm
H}=10^{5}-10^{7}$. We also consider two kinds of spatial
distribution for the radiation sources with respect to the sphere:
central and homogeneous. We find that the line width and the amount of
Lyman alpha photons escaping in the line center increase with
rotational velocity. Higher rotational velocities increase the width
of each peak in the outgoing line profile while it also increases the
amount of Lyman alpha photons escaping in the line. For large
rotational velocities and high Hydrogen optical depths the outgoing
line presents single peak a the line center. This trend is more
pronounced for radiation sources homogeneously distributed in the
sphere. Rotation increases by a factor of $2$ the escape fraction for
systems with homogeneously distributed sources. For radiation sources
located off-center, the line morphology presents a range of single,
double and triple peak lines. This work demonstrates that gas bulk
rotation has a non negligible impact on the morphology of the Lyman
$\alpha$ line.

@@

This paper is paper is structured as follows. In \S
\ref{sec:implementation} we present the implementation of bulk
rotation into the Monte Carlo codes, paying special attention to coordinate
definitions. There we also list the different physical parameters in
the simulated grid of models. In the next section we present the
results of the simulations, with special detail to quantities in the
line that show a clear evolution as a function of the cloud's
rotational velocity. In \S \ref{sec:discussion} we discuss the
implications of our results in the interpretation of LAEs
observations and further refinements that can be implemented in the
theoretical modelling of the \ly line. In the last section we present
our conclusions.

@@

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
Figure clearly shows that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $X_{\rm max}=0$ \kms.
